---
title: "make_model_code"
output: html_document
---

Purpose of script: test out pasting of model blocks dependent on presence of censored/observed subjects. Previously running into if/if else statement issues when trying to do definition time dynamic creation of nodes. Starting with just the Guassian (identity) model for simplicity. 

```{r}
## load libraries
library(nimble)
library(truncnorm)
library(coda)
library(tidyverse)
library(nimbleNoBounds)

## source helper functions
dir <- "/work/users/a/g/agfoes/P1"
source(paste0(dir, "/R/helpers/helpers_data.R"))                                        # datagen
source(paste0(dir, "/R/helpers/helpers_nimble.R"))                                      # nimble_constants, nimble_data_years, nimble_inits
source(paste0(dir, "/R/helpers/helpers_sampler.R"))                                     # make_conf, run_sampler
source(paste0(dir, "/R/helpers/helpers_summary.R"))                                     # posterior_summary, data_censoring_summary

source(paste0(dir, "/R/model_code/glm_outcome_joint_models.R"))                         # model code for each GLM distribution - gamma, poisson, negbin, bernoulli, normal

source(paste0(dir, "/R/custom_block_sampler.R"))


set.seed(08312026)
```

# create text blocks for each node we could need to add

## subject blocks
```{r}
obs_block <- "
  for (k in 1:nobs) {
    
    # cluster assignment
    xi[idx_obs[k]] ~ dcat(w[1:L])
    
    # assign component parameters based on cluster assignment
    for (j in 1:p) {
      gamma[idx_obs[k],j] <- gammaTilde[xi[idx_obs[k]],j]
    }
    sigmasq[idx_obs[k]] <- sigmasqTilde[xi[idx_obs[k]]]
    
    # latent X
    x_obs[k] ~ dnorm(
      mean = inprod(gamma[idx_obs[k], 1:p], z_obs[k, 1:p]),
      var = sigmasq[idx_obs[k]])
    
    # outcome model
    if (p > 1) {
      eta[idx_obs[k]] <- inprod(z_obs[k, 1:p], beta[2:(p+1)]) + beta[1]*x_obs[k]
    } else if (p == 1) {
      eta[idx_obs[k]] <- z_obs[k, 1] * beta[2] + beta[1] * x_obs[k]
    }
    
    
    y_obs[k] ~ dnorm(
      mean = eta[idx_obs[k]],
      tau = tau
    )
  }
"

cen_block <- "
  for (k in 1:ncen) {
    
    # cluster assignment
    xi[idx_cen[k]] ~ dcat(w[1:L])
    
    # assign component parameters based on cluster assignment
    for (j in 1:p) {
      gamma[idx_cen[k],j] <- gammaTilde[xi[idx_cen[k]],j]
    }
    sigmasq[idx_cen[k]] <- sigmasqTilde[xi[idx_cen[k]]]
    
    # latent X
    x_cen[k] ~ dnorm(
      mean = inprod(gamma[idx_cen[k], 1:p], z_cen[k, 1:p]),
      var = sigmasq[idx_cen[k]])
    
    # outcome model
    if (p > 1) {
      eta[idx_cen[k]] <- inprod(z_cen[k, 1:p], beta[2:(p+1)]) + beta[1]*x_cen[k]
    } else if (p == 1) {
      eta[idx_cen[k]] <- z_cen[k, 1] * beta[2] + beta[1] * x_cen[k]
    }
    
    
    y_cen[k] ~ dnorm(
      mean = eta[idx_cen[k]],
      tau = tau
    )
  }
"
```

## subject blocks - pz = 1
```{r}
obs_block_intercept <- "
  for (k in 1:nobs) {
    
    # cluster assignment
    xi[idx_obs[k]] ~ dcat(w[1:L])
    
    # assign component parameters based on cluster assignment
    gamma[idx_obs[k]] <- gammaTilde[xi[idx_obs[k]]]
    sigmasq[idx_obs[k]] <- sigmasqTilde[xi[idx_obs[k]]]
    
    # latent X
    x_obs[k] ~ dnorm(
      mean = gamma[idx_obs[k]] * z_obs[k],
      var = sigmasq[idx_obs[k]])
    
    # outcome model
    eta[idx_obs[k]] <- z_obs[k] * beta[2] + beta[1] * x_obs[k]
    
    
    y_obs[k] ~ dnorm(
      mean = eta[idx_obs[k]],
      tau = tau
    )
  }
"

cen_block_intercept <- "
  for (k in 1:ncen) {
    
    # cluster assignment
    xi[idx_cen[k]] ~ dcat(w[1:L])
    
    # assign component parameters based on cluster assignment
    gamma[idx_cen[k]] <- gammaTilde[xi[idx_cen[k]]]
    sigmasq[idx_cen[k]] <- sigmasqTilde[xi[idx_cen[k]]]
    
    # latent X
    x_cen[k] ~ dnorm(
      mean = gamma[idx_cen[k]] * z_cen[k],
      var = sigmasq[idx_cen[k]])
    
    # outcome model
    eta[idx_cen[k]] <- z_cen[k] * beta[2] + beta[1] * x_cen[k]
    
    
    y_cen[k] ~ dnorm(
      mean = eta[idx_cen[k]],
      tau = tau
    )
  }
"
```

## unconditional blocks
```{r}
base_block <- "
  beta[1:(p+1)] ~ dmnorm(
    mean = beta_mean[1:(p+1)],
    cov = beta_cov[1:(p+1), 1:(p+1)]
  )
  
  tau ~ dgamma(1,1)
  lalpha ~ dLogGamma(1, 1)
  alpha <- exp(lalpha)
  
  ### stick-breaking
  for (h in 1:(L-1)) {
    v[h] ~ dbeta(1, alpha)
  }
  w[1:L] <- stick_breaking(v[1:(L-1)])
  
  ### component-level parameters
  for (h in 1:L) {
    sigmasqTilde[h] ~ dinvgamma(1,1)
    
    for (j in 1:p) {
      gammaTilde[h, j] ~ dnorm(mu_gamma, var = sigmasqTilde[h])
    }
  }
"
base_block_intercept <- "
  beta[1:(p+1)] ~ dmnorm(
    mean = beta_mean[1:(p+1)],
    cov = beta_cov[1:(p+1), 1:(p+1)]
  )
  
  tau ~ dgamma(1,1)
  lalpha ~ dLogGamma(1, 1)
  alpha <- exp(lalpha)
  
  ### stick-breaking
  for (h in 1:(L-1)) {
    v[h] ~ dbeta(1, alpha)
  }
  w[1:L] <- stick_breaking(v[1:(L-1)])
  
  ### component-level parameters
  for (h in 1:L) {
    sigmasqTilde[h] ~ dinvgamma(1,1)
    gammaTilde[h] ~ dnorm(mu_gamma, var = sigmasqTilde[h])
  }
"
```

# model creation based on pasting in blocks
```{r}
obs_only <- paste("{", base_block, obs_block, "}")
model_code <- eval(call("nimbleCode", parse(text = obs_only)[[1]]))

```


# test compiled model with no censored subjects
```{r}
set.seed(09022026)
data <- datagen(alpha_width = c(10, 1, 1), alpha_start = c(1, 1, 1), eps_obs = 10e6) # creating arbitrarily large obs window to get all interval censored and then increased tolerance for "closeness" of endpoints to force all subjects observed

pz <- 3
sigma_bx <- 1
sigma_bz <- 1
mu_gamma <- 0

Nconstants <- nimble_constants(data,
                               L = 50, pz = pz,
                               sigma_bx = sigma_bx,
                               sigma_bz = sigma_bz,
                               mu_gamma = mu_gamma)
Ndata <- nimble_data_years(data,
                           constants = Nconstants)
Ninits <- nimble_inits(data,
                       pz = pz,
                       constants = Nconstants,
                       Ndata = Ndata)

## fit model and get samples
model <- nimbleModel(
  code = model_code,
  data = Ndata,
  constants = Nconstants,
  inits = Ninits
)

fit <- run_sampler(model,
                   sampler_type = "custom_joint_obs",
                   pz = pz,
                   L = 50,
                   niter = 10000,
                   nburnin = 4000,
                   thin = 1,
                   vars_to_monitor = c("beta", "tau"),
                   nobs = Nconstants$nobs,
                   ncen = Nconstants$ncen
                   )

summary(fit$samples)
```


# test compiled model with only censored subjects
```{r}
set.seed(09022026)
data <- datagen(alpha_width = c(10, 1, 1), alpha_start = c(1, 1, 1), eps_obs = 0) # creating arbitrarily large obs window to get all interval censored and then reduce tolerance for "closeness" of endpoints to force all subjects censored

cen_only <- paste("{", base_block, cen_block, "}")
model_code <- eval(call("nimbleCode", parse(text = cen_only)[[1]]))

pz <- 3
sigma_bx <- 1
sigma_bz <- 1
mu_gamma <- 0

Nconstants <- nimble_constants(data,
                               L = 50, pz = pz,
                               sigma_bx = sigma_bx,
                               sigma_bz = sigma_bz,
                               mu_gamma = mu_gamma)
Ndata <- nimble_data_years(data,
                           constants = Nconstants)
Ninits <- nimble_inits(data,
                       pz = pz,
                       constants = Nconstants,
                       Ndata = Ndata)

## fit model and get samples
model <- nimbleModel(
  code = model_code,
  data = Ndata,
  constants = Nconstants,
  inits = Ninits
)

fit <- run_sampler(model,
                   sampler_type = "custom_joint_cen",
                   pz = pz,
                   L = 50,
                   niter = 10000,
                   nburnin = 4000,
                   thin = 1,
                   vars_to_monitor = c("beta", "tau"),
                   ncen = Nconstants$ncen
                   )

summary(fit$samples)
```


# test compiled model with some censored subjects, some uncensored subjects
```{r}
set.seed(09022026)
data <- datagen(alpha_width = c(10, 1, 1), alpha_start = c(1, 1, 1), eps_obs = 400) # creating arbitrarily large obs window to get all interval censored and then reduce tolerance for "closeness" of endpoints to get around 56% censoring

mixed <- paste("{", base_block_intercept, cen_block_intercept, obs_block_intercept, "}")
model_code <- eval(call("nimbleCode", parse(text = mixed)[[1]]))

pz <- 1
sigma_bx <- 1
sigma_bz <- 1
mu_gamma <- 0

Nconstants <- nimble_constants(data,
                               L = 50, pz = pz,
                               sigma_bx = sigma_bx,
                               sigma_bz = sigma_bz,
                               mu_gamma = mu_gamma)
Ndata <- nimble_data_years(data,
                           constants = Nconstants)
Ndata[["z_obs"]] <- as.vector(Ndata[["z_obs"]][, 1])
Ndata[["z_cen"]] <- as.vector(Ndata[["z_cen"]][, 1])



Ninits <- nimble_inits(data,
                       pz = pz,
                       constants = Nconstants,
                       Ndata = Ndata)
Ninits[["gammaTilde"]] <- as.vector(Ninits[["gammaTilde"]])

## fit model and get samples
model <- nimbleModel(
  code = model_code,
  data = Ndata,
  constants = Nconstants,
  inits = Ninits
)

fit <- run_sampler(model,
                   sampler_type = "custom_joint_mixed_intercept",
                   pz = pz,
                   L = 50,
                   niter = 10000,
                   nburnin = 4000,
                   thin = 1,
                   vars_to_monitor = c("beta", "tau"),
                   ncen = Nconstants$ncen,
                   nobs = Nconstants$nobs
                   )

summary(fit$samples)
```
